Insights on the Impact of Context in the Perception of a Robot

### 2015-06-15 main.tex
JHRI template + partial update with 57 subjects + formatting

%\documentclass{acm_proc_article-sp}
\documentclass[a4,twocolumn,10pt]{article}

@@

@@

\title{New Insights on the Role of Context in the Perception of a Human-like Robot}

\author{
Séverin Lemaignan, Ashish Ranjan Jha, Kshitij Sharma, Pierre Dillenbourg\\
CHILI Lab, École Polytechnique Fédérale de Lausanne\\
\url{firstname.lastname@epfl.ch}
}

\begin{document}
\maketitle
\begin{abstract}

What leads us to attribute human-like characteristics to robots? The physical
appearance of the robot and the design of its behaviours do obviously play a key
role. This paper however evidences another mechanism at play, more subtle and partially
unconscious: the context of the interaction, and more precisely, the cognitive
prerequisites that the interaction presupposes.

@@

This second point has important implications regarding the design of human-like human-robot
interactions, and we propose to discuss them at the end of the article.

%We present a study that investigates the relation between anthropomorphism and
%gaze patterns of the observers. Anthropomorphism usually refers to attributing
%human-like characteristics by a human to a non-human agent/entity; in this
%study, the non-human entity is a robot. 
%
%We hypothesize that while observing a human-robot interaction in a particular
%setting, the observer develops an anthropomorphic behavior for the robot.
%Moreover, the resulting anthropomorphism can be correlated with the gaze
%patterns of the observer as well as with the recorded response of the
%participants about their experience of the interaction (in the form of a
%questionnaire), both before and after the experiment. We try to seek the answer
%to the question whether anthropomorphism goes beyond the shape and body
%features, and we indeed observe that just by the use of audio commands of
%different cognitive levels, anthropomorphic attitude gets affected. We obtain
%two other useful results. First, eye tracking serves to be an effective way for
%measuring anthropomorphism. Second, in our study, the standard anthropomorphism
%tests are also validated such as, a person who can anthropomorphize more, in
%general, responds to a human scene and a robot scene in a similar manner.
%
%We observe a significant increase in the gaze fixations over the actor's head
%for a high cognitive scene compared to a low cognitive scene. We also achieve
%the result that the high cognitive interaction scenarios are capable of inducing
%anthropomorphic behavior more than the low cognitive scenarios.

\end{abstract}

%\keywords{anthropomorphism, eye tracking, gaze patterns, robot}

@@

    \caption{\small The study follows a 2x2 design: the \emph{robot} vs
        \emph{human} condition is within subject, while the \emph{shallow
        cognitive context} vs \emph{deep cognitive context} is between subject.
        Two stimuli (\emph{Picking} task, {\bf A}, and \emph{Sound} task,
        {\bf B}) were shown to the participants, introduced by brief verbal
        commands, reproduced in the table. While visually identical (see
        screenshots on figure~\ref{fig:stimuli}), the stimuli were introduced
        with different commands in the two conditions \emph{shallow} and
        \emph{deep cognitive
        context}.}

@@

We did a pilot test using 10 participants first, and then after some
modifications, we did the final experiment on 40 participants.

Participants on an average were around 21 years of age (SD=4.5). Among 40
participants, 21 were females and 19 were males.

The participants were recruited amongst students of our university. To avoid students
with a higher familiarity with robots and artificial intelligent systems,
Computer Science and Electronics curriculum were excluded. This was \textit{a
posteriori} comforted as per the questionnaire responses: no
participants\fxwarning{to be checked : actually 5 out of 40 reported to be familiar(they marked +1 or +2 in questionnaire)}
reported itself as being familiar with robots and only a few\fxwarning{how
many? ( = 8 out of 40 )} out of 40 had ever owned a robot.

It took
approximately 15 minutes per participant for one experiment session.  Each
participant was invited for only one session.

The \emph{shallow} vs \emph{deep cognitive context} condition was between
subject: Out of the 40 participants, 22 watched the \emph{deep cognitive}
tasks and 18 watched the \emph{shallow cognitive} tasks.

The \emph{robot} vs \emph{human} condition was within subject. The order of
stimuli presentation was counter-balanced over the 40 participants.

@@

Before leaving, each of the participants were given a reward equivalent to CHF
10.

@@

\paragraph{Areas Of Interest(AOI)}

Among the two scenarios, the first was one was about the robot/human picking up
a toy and the second one about pointing to the noise. As seen in
figure~\ref{fig:aoi-toys}, in
the first scenario, we had 10 AOIs: 1 for the actor's (robot/human) head, 2 for
arms (left and right), 2 for hands, 2 for legs and 1 for torso, and also 2 for
the two toys (green and brown). Similarly, as shown in
figure~\ref{fig:aoi-noise}, we had 10
AOIs for the second scenario, where the 2 AOIs for toys were replaced by two
speakers that produced sound. We fetched the gaze patterns that fall into one of
these 10 AOIs and analyzed the gaze distribution among them.

@@

\paragraph{Proportion of time on the AOIs}

We fetch only the gaze patterns that fall into one of these 10 AOIs and analyze
the gaze distribution among them based on the proportion of time spent in
viewing each of these AOIs. This helps in drawing comparisons between various
objects present in the scene.

@@

An important data to be obtained here is $\delta_{H,R}$. This is calculated by
taking the difference between the ratio of net dwell time of an AOI and the AOI
fixation coverage summed over all the AOIs in the scene. Hence,
$\delta_{\mathcal{H},\mathcal{R}}$ is expressed as a ratio in the range of 0 to
1. Similarly, the differences between gaze patterns of two participants, where
one watches the LC task, given by $\delta_{low}^{\mathcal{H},\mathcal{R}}$ and
the other watches the HC task, given by
$\delta_{high}^{\mathcal{H},\mathcal{R}}$ are important gaze variables.

@@

Based on the experiments conducted on 40 participants in the $\mathcal{H}$ and
robot interactions across LC and HC conditions, we have the following
results. (The detailed statistics and code used for obtaining the results are
available here : \url{https://github.com/chili-epfl/anthropomorphism-eyetracking})

@@

In our post-questionnaire, we had a question asking the participants about
whether the video has a robot kind of task or a human kind of task on a scale of
-2 to 2 where -2 refers to robot kind of task. As can be seen in figure 2, we
got a significant correlation in this manipulation check (F[1,36] = 20.42, p <
.01) which means that the participants that were shown the LC videos identified
the task as the robot-like and participants watching HC videos perceived the
tasks as more human-like.

\subsection{H1: Gaze patterns for Human vs Robot}

Among the 10 AOIs, we grouped the AOIs corresponding to the actor into 4
categories : Head, Arm (containing both arms and hands), Torso, Leg(containing
both legs). And this is the result that we obtained from the ANOVA between gaze
patterns on the AOIs versus the human and robot videos :

\textit{\textbf{head}}: the fixations on the head for the human video
were observed to more than that for the robot video (F[1,36] = \ 6.60, p
< .05)

\textit{\textbf{arm}}\ : the fixations on the arms for the $\mathcal{H}$ video
were also greater than that for the robot video (F[1,36] = 18.65, p <
.01)

\textit{\textbf{leg}}\hspace{0.18cm} : the fixations on the legs were much lower
for the human video as compared to the robot video (F[1,36] = \
2.89, p < .1)

\textit{\textbf{torso}} : the fixations on the torso was more for the
human video than for the robot video (F[1,36] = \ 3.54, p < .1)

\subsection{H2: Difference in gaze patterns vs \anti}

@@

In our experiment, we created two kinds of scenarios each for human and
robot interaction and just by changing the audio commands (priming
effect), we replicated these videos to create a HC and a LC scene. We recorded
the gaze patterns of the 40 participants who were made to watch these videos and
also the participants were asked to fill two questionnaires before and after the
experiment. These data were analyzed and compared to our four hypotheses. We
hypothesized that one who anthropomorphizes more would have a similar reaction
to human and robot conditions, and also that in the HC case, a
participant's fixations would be more on the head of the actor as compared to
that in the LC case. We also hypothesized that the HC scene would induce more
anthropomorphic attitude than the LC scene. We found that the results were
successful and supported our hypotheses. We also had another scenario where the
robot/human was asked to move/dance, but this scenario had only one object
(\ie the actor) showing movements, and hence participants would focus
more on seeing the movements leaving behind the cognition aspect, which made it
difficult to differentiate between the LC and HC cases and, therefore, we
discarded this scenario from our analysis.

@@

\bibliographystyle{abbrv}
\bibliography{biblio}

>>>>

\documentclass[lettersize, apacite, twoside, HRI]{apa_HRI}

@@

\usepackage{times}

@@

\rightheader{Impact of Context in the Perception of  a Robot}   % This should be the title, or a shortened version of the title if your title is long.
\leftheader{Lemaignan et al.}	  % For one or two authors, include both authors last names.  For three or more, use first author's last name et al.
\title{Insights on the Impact of Context in the Perception of a Robot}

\author{Séverin Lemaignan, Ashish Ranjan Jha, Kshitij Sharma, Pierre
Dillenbourg}
\affiliation{CHILI Lab, École Polytechnique Fédérale de Lausanne}

\acknowledgements{\url{firstname.lastname@epfl.ch}}

\abstract{
What leads us to attribute human-like characteristics to robots? The physical
appearance of the robot and the design of its behaviours do obviously play a key
role. This paper however evidences another mechanism at play, more subtle and partially
unconscious: the context of the interaction, and more precisely, the cognitive
prerequisites that the interaction presupposes.

@@

This second point has important implications regarding the design of human-like human-robot
interactions, and we propose to discuss them at the end of the article.
}

\keywords{human-robot interaction, anthropomorphism, cognitive context, eye tracking, gaze patterns}

\begin{document}
\maketitle

@@

    \caption{\small The study follows a 2$\times$2 design: the \emph{robot} vs
        \emph{human} condition is within subject, while the \emph{shallow
        cognitive context} vs \emph{deep cognitive context} is between subject.
        Two stimuli (\emph{Picking} task, {\bf A}, and \emph{Sound} task,
        {\bf B}) were shown to the participants, introduced by brief verbal
        commands, reproduced in the table. While visually identical (see
        screenshots on figure~\ref{fig:stimuli}), the stimuli were introduced
        with different commands in the two conditions \emph{shallow} and
        \emph{deep cognitive
        context}.}

@@

We ran first a pilot study with 10 participants, followed by an experiment with
57 subjects (mean age: 20.9, $SD=3.9$, 32 females and 25 males).

The participants were recruited amongst students of our university. To avoid
students with a high familiarity with robots and artificial intelligent systems,
Computer Science and Electronics curriculum were excluded. This was \textit{a
posteriori} comforted as per the questionnaire responses: no participants
reported itself as being very familiar with robots, 8 reported some familiarity
and 7 out of 57 own a domestic robot (vacuum cleaner).

The study lasted about 15 minutes per participant; each participant was invited
for one session only. Before leaving, each participant was given a reward
equivalent to EUR 10.

The study follows a 2$\times$2 design (table~\ref{table:design}): the
\emph{deep} vs {shallow cognitive context} condition was between-subject ( 31
subjects performed the \emph{deep cognitive} tasks while 26 were in the
\emph{shallow cognitive} condition); the \emph{robot} vs \emph{human} condition
was within-subject (the order of stimuli presentation was counter-balanced over
the 57 participants).

@@

@@

\paragraph{Areas Of interest}

Among the two scenarios, the first was one was about the robot/human picking up
a toy and the second one about pointing to the noise. As seen in
figure~\ref{fig:aoi}, we defined 10 areas of interest (AOIs): one on the head of
the agent (robot or human), two on the arms (left and right), two on the hands,
two on the legs and one on the torso. Besides, we defined one area of interest
per toy or speaker (depending on the stimulus).

@@

We collected the gazes falling into one of these 10 zones and analyzed
their distribution (producing what we call \emph{gaze patterns}).

@@

An important data to be obtained here is $\delta_{\mathcal{H},\mathcal{R}}$. This is calculated by
taking the difference between the ratio of net dwell time on an AOI and the AOI
fixation coverage summed over all the AOIs in the scene. Hence,
$\delta_{\mathcal{H},\mathcal{R}}$ is expressed as a ratio in the range of 0 to
1. Similarly, the differences between gaze patterns of two participants, where
one watches the LC task, given by $\delta_{shallow}^{\mathcal{H},\mathcal{R}}$ and
the other watches the HC task, given by
$\delta_{deep}^{\mathcal{H},\mathcal{R}}$ are important gaze variables.

@@

Based on the experiments conducted on 57 participants in the human
($\mathcal{H}$) and robot ($\mathcal{R}$) interactions across \emph{shallow} and
\emph{deep cognitive context} conditions, we have the following results. (the
detailed statistics and code used for obtaining the results are available here :
\url{https://github.com/chili-epfl/anthropomorphism-eyetracking})

@@

In our post-questionnaire, we had a question asking the participants about
whether the video has a robot kind of task or a human kind of task on a scale of
-2 to 2 where -2 refers to robot kind of task. As can be seen in figure 2, we
got a significant correlation in this manipulation check ($F[1,36] = 20.42, p <
.01$) which means that the participants that were shown the \emph{shallow
cognitive context} videos identified the task as more robot-like and
participants watching \emph{deep cognitive context} videos perceived the tasks
as more human-like.

\subsection{H1: Gaze differences between human and robot conditions}

We grouped the 8 areas of interest located on the agent into 4 categories : {\sf
head}, {\sf arm} (containing both arms and hands), {\sf torso}, {\sf leg}
(including both legs). And this is the result that we obtained from the ANOVA
between gaze patterns on the AOIs versus the human and robot videos :

{\sf head}: the fixations on the head for the human video
were observed to more than that for the robot video ($F[1,36] = 6.60, p < .05$)

{\sf arm}: the fixations on the arms for the $\mathcal{H}$ video
were also greater than that for the robot video ($F[1,36] = 18.65, p < .01$)

{\sf leg}: the fixations on the legs were much lower
for the human video as compared to the robot video ($F[1,36] = 2.89, p < .1$)

{\sf torso}: the fixations on the torso was more for the
human video than for the robot video ($F[1,36] = 3.54, p < .1$)

\subsection{H2: Interaction between \anti and gaze patterns}

@@

In our experiment, we created two kinds of scenarios each for human and
robot interaction and just by changing the audio commands (priming
effect), we replicated these videos to create a HC and a LC scene. We recorded
the gaze patterns of the 57 participants who were made to watch these videos and
also the participants were asked to fill two questionnaires before and after the
experiment. These data were analyzed and compared to our four hypotheses. We
hypothesized that one who anthropomorphizes more would have a similar reaction
to human and robot conditions, and also that in the HC case, a
participant's fixations would be more on the head of the actor as compared to
that in the LC case. We also hypothesized that the HC scene would induce more
anthropomorphic attitude than the LC scene. We found that the results were
successful and supported our hypotheses. We also had another scenario where the
robot/human was asked to move/dance, but this scenario had only one object
(\ie the actor) showing movements, and hence participants would focus
more on seeing the movements leaving behind the cognition aspect, which made it
difficult to differentiate between the LC and HC cases and, therefore, we
discarded this scenario from our analysis.

@@

\bibliography{biblio}
